### 2018-07-08 figures/ensemble_scheme.tex

\frame{\begin{picture}(100,100)
\put(25,50){\dashbox{.5}(10,10){$C$}}
\put(50,50){\dashbox{.5}(10,10){$B$}}
\end{picture}}